---
title: "SRT411A0"
author: "Connor Brozic"
date: "February 1, 2017"
output:
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Intro
Hello, this is Assignment 0 for SRT411.  This was written by Connor Brozic.

Github Username: ConnorBrozic

This assigment focuses on completing the 15 "ToDos" from within the "A (very) short introduction to R"", written by Paul Torfs and Claudia Brauer.  It focuses on the basics of the R languages, to showcase how powerful even a basic part of R is.  
This document can be found at:  https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf

I have used help sources in some cases, which i have outlined at the bottom of this document.  

#### ToDo Question 1
Compute the difference between 2014 and the
year you started at this university and divide
this by the difference between 2014 and the year
you were born. Multiply this with 100 to get
the percentage of your life you have spent at
this university.
```{r}
#Modifications changed to keep dates current
((2017-2015)/(2017-1997))*100
```

#### ToDo Question 2
Repeat the previous ToDo, but with several
steps in between. You can give the variables
any name you want, but the name has to start
with a letter.
```{r}
CurrentMinusStarting <- 2017-2015
CurrentMinusBirth <- 2017-1997
DecimalofLife <- CurrentMinusStarting/CurrentMinusBirth
PercentofLife <- DecimalofLife * 100
PercentofLife
```

#### ToDo Question 3
Compute the sum of 4, 5, 8 and 11 by first combining
them into a vector and then using the
function sum.
```{r}
sum(ToDo3 <- c(4,5,8,11))
```

#### ToDo Question 4
Plot 100 normal random numbers.
```{r}
ToDo4 <- rnorm(100)
plot(ToDo4)
```

#### ToDo Question 5
Find help for the sqrt function.
```{r}
help(sqrt)
```

#### ToDo Question 6
Make a file called firstscript.R containing Rcode
that generates 100 random numbers and
plots them.
```{r}
source("firstscript.R")

#Code within the firstscript.R file.
#ToDo6 <- rnorm(100)
#plot(ToDo6)
```

#### ToDo Question 7
Put the numbers 31 to 60 in a vector named
P and in a matrix with 6 rows and 5 columns
named Q.
```{r}
P <- c(31:60)
Q <- matrix(data=P, nrow = 6, ncol = 5)
Q
```

#### ToDo Question 8
Make a script file which constructs three random
normal vectors of length 100. Call these
vectors x1, x2 and x3. Make a data frame called
t with three columns (called a, b and c) containing
respectively x1, x1+x2 and x1+x2+x3.
Call the following functions for this data frame:
plot(t) and sd(t).
```{r}
x1 <- c(rnorm(100))
x2 <- c(rnorm(100))
x3 <- c(rnorm(100))
ToDo8DataFrame = data.frame(a = x1, b = x1+x2, c = x1+x2+x3)
plot(ToDo8DataFrame) #Works no problem and plots the results.
#sd(t) #Does not work as lists cannot be converted to double. 

```

#### ToDo Question 9
Add these lines to the script file of the previous
section. Try to find out, either by experimenting
or by using the help, what the meaning is of
rgb, the last argument of rgb, lwd, pch, cex.
```{r}
plot(ToDo8DataFrame$a, type = "l", ylim = range(ToDo8DataFrame), lwd = 3, col=rgb(1,0,0,0.3))
lines(ToDo8DataFrame$b, type = "s", lwd = 2, col = rgb(0.3,0.4,0.3,0.9))
points(ToDo8DataFrame$c, pch = 20, cex = 4, col = rgb(0,0,1,0.3))
#Answer to RGB Question:  the RGB argument specifies which colour to use for the specified plot
#Takes four numbers, Red, green, blue and alpha. 
#Alpha controls opacity of the element
```
#### ToDo Question 10
Make a file called tst1.txt in Notepad from
the example in Figure 4 and store it in your
working directory. Write a script to read it, to
multiply the column called g by 5 and to store
it as tst2.txt.
```{r}
ToDo11 <- read.table(file="tst1.txt", header = TRUE)
ToDo11$g <- ToDo11$g * 5
ToDo11
write.table(ToDo11,file = "tst2.txt", row.names = TRUE)
```

#### ToDo Question 11
Compute the mean of the square root of a vector
of 100 random numbers. What happens?
```{r}
ToDo10 <- mean(sqrt(c(rnorm(100))))
#This calculation does not work, results in NaNs being produced as an output.  
```

#### ToDo Question 12
Make a graph with on the x-axis: today, Sinterklaas
2014 and your next birthday and on
the y-axis the number of presents you expect on
each of these days. 
```{r}
xAxis <- strptime(c("20170128", "20170416", "20171206"), format = "%Y%m%d")
yAxis <- c(0, 7,15)
plot(xAxis, yAxis, type = "b", col="Red", ylab="# of presents", xlab ="Date", main = "Presents Today, Birthday and Sinterklaas")

```

#### ToDo Question 13
Make a vector from 1 to 100. Make a for-loop
which runs through the whole vector. Multiply
the elements which are smaller than 5 and larger
than 90 with 10 and the other elements with 0.1.
```{r}
ToDo13 <- seq(from = 1, to = 100)
ToDo13a <- c()
for(i in ToDo13)
  {
  if (i < 5)
  {
    ToDo13a[i]= ToDo13[i] * 10
  }else if (i > 90){
    ToDo13a[i]= ToDo13[i] * 10
  }else{
    ToDo13a[i] = ToDo13[i] * 0.1
  }
}
ToDo13a
```

#### ToDo Question 14
Write a function for the previous ToDo, so
that you can feed it any vector you like
(as argument). Use a for-loop in the function
to do the computation with each element.
Use the standard R function length
in the specification of the counter
```{r}
ToDo14Function <- function(arg1)
  {
  arg1length <<- length(arg1)
  for(i in 1:arg1length)
    {
    if (arg1[i] < 5)
      {
      arg1[i]= arg1[i] * 10
      }
    else if (arg1[i] > 90){
      arg1[i]= arg1[i] * 10
    }else{
      arg1[i] = arg1[i] * 0.1
    }
  }
  return(arg1)
}
test <- c(1:100)
ToDo14Function(arg1 = test)
```

#### ToDo Question 15
The ToDo above can be done more easily
and quickly without a for-loop but with regular vector-computations.
```{r}
ToDo15 <- c(1:100)
Result1 <- ToDo15[ToDo15<5] * 10
Result2 <- ToDo15[ToDo15>=5 & ToDo15<=90] * 0.1
Result3 <- ToDo15[ToDo15>90] * 10
ToDo15 <- c(Result1,Result2,Result3)
ToDo15

```
#### Sources Used
* http://rmarkdown.rstudio.com/
* https://nicercode.github.io/guides/reports/
* http://kbroman.org/knitr_knutshell/pages/markdown.html
* http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
* https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
* https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf
* https://therostrumblog.wordpress.com/2014/01/29/basic-data-frame-manipulations-in-r/
* http://stackoverflow.com/questions/2564258/plot-two-graphs-in-same-plot-in-r
* https://stat.ethz.ch/R-manual/R-devel/library/graphics/html/axis.POSIXct.html
* http://stackoverflow.com/questions/1236620/global-variables-in-r